PyPy's Approach to \\ Implementing Dynamic Languages \\ Using a Tracing JIT Compiler

### 2014-10-01 talk/microsoft-jan-2011/talk.tex
More slides

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{ulem}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{alltt}

\newcommand\redsout[1]{{\color{red}\sout{\hbox{\color{black}{#1}}}}}

@@

\date{
%\includegraphics[scale=0.25]{figures/hpi_logo_kl.jpg}\\
IBM Watson Research Center, February 8th, 2011
}

@@

\begin{frame}
  \frametitle{How do Interpreters Meet the Requirements?}
  \includegraphics[scale=0.6]{figures/output1.pdf}
%  \begin{enumerate}
%      \item \alert{control flow slowed by bytecode or AST dispatch overhead}
%      \item \alert{late binding works but slow}
%      \item \alert{dispatching works but slow}
%      \item \alert{no special dependencies support}
%      \item \alert{everything is boxed}
%      \item \alert{reified frames are easy but slow}
%%      \item \alert{access to native libraries easy but slow}
%  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{How do Static Compilers Meet the Requirements?}
  \includegraphics[scale=0.6]{figures/output2.pdf}
%  \begin{enumerate}
%      \item control flow works well
%      \item \alert{late binding not improved}
%      \item \alert{dispatching not improved}
%      \item \alert{dependencies not improved}
%      \item \alert{everything is boxed}
%      \item \alert{reified frames often not supported}
%%      \item direct access to native libraries
%  \end{enumerate}
\end{frame}

@@

\begin{frame}[containsverbatim, plain]
  \frametitle{Method-Based JITs and Dispatching Dependencies}
  \includegraphics[scale=0.5]{figures/add1.pdf}
%\begin{verbatim}
%    if (isinstance(a, Integer) and
%            isinstance(b, Integer)):
%        x = Integer(<perform int addition>)
%    else:
%        x = Float(<perform float addition>)
%
%    # x can be Float or Integer here
%
%    if (isinstance(x, Integer) and
%            isinstance(c, Integer)):
%        r = Integer(<perform int addition>)
%    else:
%        r = Float(<perform float addition>)
%\end{verbatim}
\end{frame}

\begin{frame}[containsverbatim, plain]
  \frametitle{Method-Based JITs and Dispatching Dependencies}
  \includegraphics[scale=0.39]{figures/add2.pdf}
\end{frame}

\begin{frame}[containsverbatim, plain]
  \frametitle{Method-Based JITs and Dispatching Dependencies}
  \includegraphics[scale=0.41]{figures/add_fast.pdf}
\end{frame}

\begin{frame}
  \frametitle{How do Method-Based JIT Compilers Meet the Requirements?}
  \includegraphics[scale=0.6]{figures/output3.pdf}
%  \begin{enumerate}
%      \item control flow works well
%      \item late binding can be handled
%      \item dispatching can be handled
%      \item \alert{dependencies not necessarily improved}
%      \item boxing hard to support
%      \item \alert{reified frames hard to support}
%%      \item \alert{access to native libraries not improved}
%  \end{enumerate}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Example: Dispatching in a Tracing JIT}
\begin{verbatim}
x = ADD(a : Integer, b : Integer)
\end{verbatim}
\end{frame}

\begin{frame}[containsverbatim]
  \frametitle{Example: Dispatching in a Tracing JIT}
\begin{verbatim}
guard_class(a, Integer)
guard_class(b, Integer)
u_a = unbox(a)
u_b = unbox(b)
u_x = int_add(a, b)
x = new(Integer, u_x)
\end{verbatim}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Example: Dependencies in a Tracing JIT}
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
u_a = unbox(a)
u_b = unbox(b)
u_x = int_add(u_a, u_b)
x = new(Integer, u_x)

guard_class(x, Integer)
guard_class(c, Integer)
u_x2 = unbox(x)
u_c = unbox(c)
u_r = int_add(u_x2, u_c)
r = new(Integer, u_r)
\end{alltt}
\end{frame}

\begin{frame}[containsverbatim]
  \frametitle{Example: Dependencies in a Tracing JIT}
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
u_a = unbox(a)
u_b = unbox(b)
u_x = int_add(u_a, u_b)
x = new(Integer, u_x)

\redsout{guard_class(x, Integer)}
guard_class(c, Integer)
u_x2 = unbox(x)
u_c = unbox(c)
u_r = int_add(u_x2, u_c)
r = new(Integer, u_r)
\end{alltt}
\end{frame}

@@

\begin{frame}[containsverbatim]
  \frametitle{Example: Boxing in a Tracing JIT}
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
u_a = unbox(a)
u_b = unbox(b)
u_x = int_add(u_a, u_b)
x = new(Integer, u_x)

guard_class(c, Integer)
u_x2 = unbox(x)
u_c = unbox(c)
u_r = int_add(u_x2, u_c)
r = new(Integer, u_r)
\end{alltt}
\end{frame}

\begin{frame}[containsverbatim]
  \frametitle{Example: Boxing in a Tracing JIT}
\begin{alltt}
guard_class(a, Integer)
guard_class(b, Integer)
u_a = unbox(a)
u_b = unbox(b)
u_x = int_add(u_a, u_b)
\redsout{x = new(Integer, u_x)}

guard_class(c, Integer)
\redsout{u_x2 = unbox(x)}
u_c = unbox(c)
u_r = int_add(u_x, u_c)
r = new(Integer, u_r)
\end{alltt}
\end{frame}

@@

\begin{frame}
  \frametitle{How do Tracing JITs Meet the Requirements?}
  \includegraphics[scale=0.6]{figures/output4.pdf}
%  \begin{enumerate}
%      \item control flow works rather well
%      \item late binding can be handled
%      \item dispatching can be handled
%      \item dependencies improved by path splitting
%      \item unboxing optimization much simpler
%      \item reified frames can be implemented \\
%      by falling back to the interpreter
%%      \item \alert{access to native libraries not improved}
%  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{How do OO VMs Meet the Requirements?}
  \includegraphics[scale=0.6]{figures/output5.pdf}
%  \begin{enumerate}
%      \item control flow works well
%      \item late binding can be handled with a lot of effort
%      \item dispatching can be handled with a lot of effort
%      \item \alert{dependencies not improved}
%      \item \alert{boxing not necessarily improved}
%      \item \alert{reified frames are inefficient}
%%      \item access to native libraries not improved, but VM libraries help
%  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{Project Status}
  \begin{itemize}
      \item about two people work on it full-time
      \item sizeable open source community
      \item one-week development sprints several times per year
      \item about 20-30 person-years so far
      \item heavily dedicated to testing and quality
  \end{itemize}
\end{frame}

\begin{frame}[plain]
  \includegraphics[scale=0.25]{figures/timeline.png}
\end{frame}

@@

\begin{frame}[plain]
  \frametitle{Architecture}
  \includegraphics[scale=0.4]{figures/architecture.pdf}
\end{frame}

>>>>

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{ulem}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{alltt}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning,shadows,shapes,calc,backgrounds,fit}

@@

\date{
%\includegraphics[scale=0.25]{figures/hpi_logo_kl.jpg}\\
%IBM Watson Research Center, February 9th, 2011
Microsoft Research, January 31st, 2011
}

@@

\begin{frame}
  \frametitle{How do Interpreters Meet the Requirements?}
  \begin{enumerate}
      \item \alert{control flow slowed by bytecode or AST dispatch overhead}
      \item \alert{late binding works but slow}
      \item \alert{dispatching works but slow}
      \item \alert{no special dependencies support}
      \item \alert{everything is boxed}
      \item \alert{reified frames are easy but slow}
%      \item \alert{access to native libraries easy but slow}
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{How do Static Compilers Meet the Requirements?}
  \begin{enumerate}
      \item control flow works well
      \item \alert{late binding not improved}
      \item \alert{dispatching not improved}
      \item \alert{dependencies not improved}
      \item \alert{everything is boxed}
      \item \alert{reified frames often not supported}
%      \item direct access to native libraries
  \end{enumerate}
\end{frame}

@@

\begin{frame}[containsverbatim, plain]
  \frametitle{Method-Based JITs and Dispatching Dependencies}
\begin{verbatim}
    if isinstance(a, Integer):
        if isinstance(b, Integer):
            x = Integer(<perform int addition>)
        else:
            x = Float(<perform float addition>)
    else:
        x = Float(<perform float addition>)
    # x can be Float or Integer here
    if isinstance(x, Integer):
        if isinstance(c, Integer):
            r = Integer(<perform int addition>)
        else:
            r = Float(<perform float addition>)
    else:
        r = Float(<perform float addition>)
\end{verbatim}
\end{frame}

\begin{frame}
  \frametitle{How do Method-Based JIT Compilers Meet the Requirements?}
  \begin{enumerate}
      \item control flow works well
      \item late binding can be handled
      \item dispatching can be handled
      \item \alert{dependencies not necessarily improved}
      \item boxing hard to support
      \item \alert{reified frames hard to support}
%      \item \alert{access to native libraries not improved}
  \end{enumerate}
\end{frame}

@@

@@

@@

@@

\begin{frame}
  \frametitle{How do Tracing JITs Meet the Requirements?}
  \begin{enumerate}
      \item control flow works rather well
      \item late binding can be handled
      \item dispatching can be handled
      \item dependencies improved by path splitting
      \item unboxing optimization much simpler
      \item reified frames can be implemented \\
      by falling back to the interpreter
%      \item \alert{access to native libraries not improved}
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{How do OO VMs Meet the Requirements?}
  \begin{enumerate}
      \item control flow works well
      \item late binding can be handled with a lot of effort
      \item dispatching can be handled with a lot of effort
      \item \alert{dependencies not improved}
      \item \alert{boxing not necessarily improved}
      \item \alert{reified frames are inefficient}
%      \item access to native libraries not improved, but VM libraries help
  \end{enumerate}
\end{frame}

@@

@@
